---
title: "Regressions for Macroinvertebrate Responses"
author: "James Orr"
output:
   html_notebook:
     code_folding: hide
     theme: flatly
     toc: true
     toc_depth: 4
     number_sections: no
---


## Introduction

In this notebook I do a small amount of organising and exploring of the combined water quality and invert data. I then run linear mixed effects models to create some "offline" models that we can use to make predictions using the INCA model scenario outcomes. 

Response variables that can be predicted by INCA are: 

- Flow (m3/s)

- Soluble Reactive Phosphorus (mg/l)

- Chlorophyll a (mg/l)

- Temperature (degrees)

- Dissolved Oxygen (mg/l)

- Nitrate (NO3)

- Ammonium (NH4) 

Phosphorus, dissolved oxygen, temperature and nitrate are all be directly incorporated into my regressions using EA data. Flow can be estimated and ammonium can be approximated using ammonia. Chlorophyll a can't be included because of insufficient data in the EA water quality dataset. 

We can also include some strong predictor variables that aren't in INCA (e.g., describing habitat quality). We will also include random effect for site to account for repeated measures across years. 

## Data preparation

**Set up environment**

```{r, echo=TRUE, results='hide'}
#### Clear  environment 
rm(list = ls())         

#### Required packages
library(tidyverse)        # for general use
library(GGally)           # for ggpairs
library(lme4)             # for lmer() function
library(MuMIn)            # to get r squared for lme4 models
library(lmerTest)         # to get p values for lme4 models
library(DHARMa)           # for diagnostics of glmm models 

```

**Load data**

```{r}
glmm_data <- list.files(path = "data/combined/",
                  pattern="*.csv",
                  full.names = T) %>%
  map_df(~read_csv(., col_types = cols(.default = "c"))) 
```

**Select the variables we're broadly interested in**

```{r}
glmm_data <- glmm_data %>%
  
  dplyr::select(c(
    
          # general descriptors of each sample 
          SITE_ID, SAMPLE_DATE_YEAR, SAMPLE_DATE_MONTH,
          
          # habitat quality variables
          DISCHARGE, WIDTH, SLOPE, 
          SILT_CLAY, BOULDERS_COBBLES, PEBBLES_GRAVEL, SAND,
          
          # water quality variables (that can be predicted by INCA)
          `Temperature of Water_cel`,
          `Oxygen, Dissolved as O2_mg/l`,
          `Ammonia un-ionised as N_mg/l`, 
          `Nitrate as N_mg/l`,
          `Orthophosphate, reactive as P_mg/l`, 
           
          # response variables 
          WHPT_ASPT, WHPT_TOTAL, 
          macro_richness, 
          mollusc, 
          total_abundance,
          ept_richness,
          ept_abundance,
          `insect - true fly (Diptera)`,
          `insect - stonefly (Plecoptera)`,
          `insect - mayfly (Ephemeroptera)`,
          `insect - dragonfly (Odonata)`,
          `insect - caddis fly (Trichoptera)`,
          `insect - beetle (Coleoptera)`,
          `flatworm (Turbellaria)`,
          crustacean,
          annelid)) %>%
  
  drop_na()  # llmer requires full cases 


# convert all to numeric 
glmm_data <- data.frame(lapply(glmm_data, function(x) as.numeric(as.character(x))))

# set site as a factor 
glmm_data$SITE_ID <- as.factor(glmm_data$SITE_ID) 

```

**Count observations per site**

```{r}
#count <- glmm_data %>%
#  count(SITE_ID)
#glmm_data <- left_join(glmm_data, count, by = "SITE_ID")
#table(count$n)
```

**Create a variable for flow (based on EA discharge classes)**

```{r}
# discharge categories are from 1 to 10 and are based on a log scale of.... 
# naturalized mean annual discharge (cumecs). For each discharge category ....
# set the flow to the middle of the range of cumec data.

glmm_data$flow <- if_else(glmm_data$DISCHARGE == 1, 0.155,
                  if_else(glmm_data$DISCHARGE == 2, 0.465,
                  if_else(glmm_data$DISCHARGE == 3, 0.935,
                  if_else(glmm_data$DISCHARGE == 4, 1.875,
                  if_else(glmm_data$DISCHARGE == 5, 3.75,
                  if_else(glmm_data$DISCHARGE == 6, 7.5,
                  if_else(glmm_data$DISCHARGE == 7, 15.0,
                  if_else(glmm_data$DISCHARGE == 8, 30.0,
                  if_else(glmm_data$DISCHARGE == 9, 60.0,
                  if_else(glmm_data$DISCHARGE == 10, 120.0, 0
                  ))))))))))


```



## Data exploration

**Co-linearity of explanatory variables**

```{r, message=F, warning=F}

hq = c(16, 4, 5, 6, 7, 8, 9, 10)

ggpairs(glmm_data[, hq], 
        lower = list(continuous = wrap("points", col = rgb(0.2, 0.5, 1, 0.4), 
                                       alpha = 0.25, size=0.3))) + 
  theme_minimal(base_size = 6)
```


```{r, message=F, warning=F}

wq = c(16, 11, 12, 13, 14, 15)

ggpairs(glmm_data[, wq], 
        lower = list(continuous = wrap("points", col = rgb(0.2, 0.5, 1, 0.4), 
                                       alpha = 0.25, size=0.3))) + 
  theme_minimal(base_size = 6)
```

## Regression models

### LMMs

**WHPT_ASPT**

```{r}
# log transform flow, ammonia, nitrate and orthophosphate as these are skewed 

### Variables that are not considered in the regressions:
# WIDTH - co-linear with flow
# SLOPE - no hypothesis 
# BOULDERS_COBBLES - not independent of SILT_CLAY
# PEBBLES_GRAVEL - not independent of SILT_CLAY
# SAND - not independent of SILT_CLAY
# poly(Temperature.of.Water_cel, 2) - no curvature and doesn't improve model
# SAMPLE_DATE_MONTH - not significant and no clear hypothesis
# SAMPLE_DATE_YEAR - not significant and no clear hypothesis
# Interactions were not considered a priori and didn't significantly increase fit

aspt_model <- lmer(WHPT_ASPT ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
              data = glmm_data)

options(scipen=999)
summary(aspt_model)

r.squaredGLMM(aspt_model)
```

**WHPT_TOTAL**

```{r}
total_model <- lmer(WHPT_TOTAL ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
              data = glmm_data)

summary(total_model)

r.squaredGLMM(total_model)
```


**Macroinvertebrate Richness**

```{r}
rich_model <- lmer(macro_richness ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
              data = glmm_data)

summary(rich_model)

r.squaredGLMM(rich_model)
```




### Diagnostics 

**1. Assess co-linearity of predictors** 

```{r}
predictors <- dplyr::select(glmm_data, c(flow, SILT_CLAY, 
                                  Temperature.of.Water_cel, 
                                  Oxygen..Dissolved.as.O2_mg.l,
                                  Orthophosphate..reactive.as.P_mg.l,
                                  Ammonia.un.ionised.as.N_mg.l,
                                  Nitrate.as.N_mg.l))

predictors$flow <- log(predictors$flow)
predictors$Ammonia.un.ionised.as.N_mg.l <- log(predictors$Ammonia.un.ionised.as.N_mg.l)
predictors$Orthophosphate..reactive.as.P_mg.l <- log(predictors$Orthophosphate..reactive.as.P_mg.l)
predictors$Nitrate.as.N_mg.l <- log(predictors$Nitrate.as.N_mg.l)

ggpairs(predictors, 
        lower = list(continuous = wrap("points", col = rgb(0.2, 0.5, 1, 0.4), 
                                       alpha = 0.25, size=0.3))) + 
  theme_minimal(base_size = 6);
```


**2. Residuals are normally distributed**

```{r}
par(mfrow=c(2,2))
hist(resid(aspt_model), breaks = 30)
hist(resid(total_model), breaks = 30)
hist(resid(rich_model), breaks = 30)

```


**3. Homogeneity of variance** 

```{r}
plot(aspt_model)
plot(total_model)
plot(rich_model)
```

### GLMMs

Taxa are zero inflated and/or heavily skewed - use Poisson distribution with glmer function.


**Plecoptera Abundance**

```{r}
plecoptera_model <- glmer(glmm_data$insect...stonefly..Plecoptera.~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
                #### error distribution #### 
                family = "poisson",
                
                data = glmm_data,
                
                # to help with model convergence
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(plecoptera_model)

r.squaredGLMM(plecoptera_model)


```


**Annelida Abundance**

```{r}

annelid_model <- glmer(glmm_data$annelid ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
                #### error distribution #### 
                family = "poisson",
                
                data = glmm_data,
                
                # to help with model convergence
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(annelid_model)

r.squaredGLMM(annelid_model)


```

**Trichoptera Abundance**

```{r}

trichoptera_model <- glmer(glmm_data$insect...caddis.fly..Trichoptera. ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
                #### error distribution #### 
                family = "poisson",
                
                data = glmm_data,
                
                # to help with model convergence
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(trichoptera_model)

r.squaredGLMM(trichoptera_model)


```



**Ephemeroptera Abundance**

```{r}

mayfly_model <- glmer(glmm_data$insect...mayfly..Ephemeroptera. ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
                #### error distribution #### 
                family = "poisson",
                
                data = glmm_data,
                
                # to help with model convergence
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(mayfly_model)

r.squaredGLMM(mayfly_model)


```

**Odonata Abundance**

```{r}

dragon_model <- glmer(glmm_data$insect...dragonfly..Odonata. ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
                #### error distribution #### 
                family = "poisson",
                
                data = glmm_data,
                
                # to help with model convergence
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(dragon_model)

r.squaredGLMM(dragon_model)


```

**Coleoptera Abundance**

```{r}

beet_model <- glmer(glmm_data$insect...beetle..Coleoptera. ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
                #### error distribution #### 
                family = "poisson",
                
                data = glmm_data,
                
                # to help with model convergence
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(beet_model)

r.squaredGLMM(beet_model)


```

**Flatworm Abundance**

```{r}

flat_model <- glmer(glmm_data$flatworm..Turbellaria. ~ 
                ######## physical variables #########
                log(flow) +  
                SILT_CLAY +
                
                ######## water quality variables #########
                Temperature.of.Water_cel +           
                Oxygen..Dissolved.as.O2_mg.l + 
                log(Orthophosphate..reactive.as.P_mg.l) +    
                log(`Nitrate.as.N_mg.l`) +   
                log(`Ammonia.un.ionised.as.N_mg.l`) +
                  
                ######## random effect of site  #########
                (1 | SITE_ID),
          
                #### error distribution #### 
                family = "poisson",
                
                data = glmm_data,
                
                # to help with model convergence
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(flat_model)

r.squaredGLMM(flat_model)


```


### Diagnostics

```{r}

# only consider groups whose regression models are above 0.1. This includes:
# flat_model, beet_model, dragon_model, mayfly_model, trichoptera_model, plecoptera_model, annelid_model

# excluded groups due to poorly fitting regressions = Diptera, Mollusca, Crustacean


### Obtain informative versions of residuals###
simulationOutput_stone <- simulateResiduals(fittedModel = plecoptera_model, 
                                            plot = F)
simulationOutput_worm <- simulateResiduals(fittedModel = annelid_model, 
                                            plot = F)
simulationOutput_caddis <- simulateResiduals(fittedModel = trichoptera_model, 
                                            plot = F)
simulationOutput_flat <- simulateResiduals(fittedModel = flat_model, 
                                            plot = F)
simulationOutput_beet <- simulateResiduals(fittedModel = beet_model, 
                                            plot = F)
simulationOutput_dragon <- simulateResiduals(fittedModel = dragon_model, 
                                            plot = F)
simulationOutput_may <- simulateResiduals(fittedModel = mayfly_model, 
                                            plot = F)


# check normality of residules (specific tests too sensitive to large n)
par(mfrow=c(2,4))
plotQQunif(simulationOutput_stone, 
           testUniformity = F, 
           testOutliers = F,
           testDispersion = F,
           main = "Plecoptera",
           pch = 19, cex = 0.2) 
plotQQunif(simulationOutput_worm, 
           testUniformity = F, 
           testOutliers = F,
           testDispersion = F,
           main = "Annelida",
           pch = 19, cex = 0.2) 
plotQQunif(simulationOutput_caddis, 
           testUniformity = F, 
           testOutliers = F,
           testDispersion = F,
           main = "Trichoptera",
           pch = 19, cex = 0.2) 
plotQQunif(simulationOutput_flat, 
           testUniformity = F, 
           testOutliers = F,
           testDispersion = F,
           main = "Flatworms",
           pch = 19, cex = 0.2) 
plotQQunif(simulationOutput_beet, 
           testUniformity = F, 
           testOutliers = F,
           testDispersion = F,
           main = "Coleoptera",
           pch = 19, cex = 0.2)
plotQQunif(simulationOutput_dragon, 
           testUniformity = F, 
           testOutliers = F,
           testDispersion = F,
           main = "Odonata",
           pch = 19, cex = 0.2) 
plotQQunif(simulationOutput_may, 
           testUniformity = F, 
           testOutliers = F,
           testDispersion = F,
           main = "Ephemeroptera",
           pch = 19, cex = 0.2) 


# check residules v predicted (specific tests too sensitive to large n)
par(mfrow=c(2,4))
plotResiduals(simulationOutput_stone, quantreg = F, pch = 19, cex = 0.2)#, main = "Plecoptera")
plotResiduals(simulationOutput_worm, quantreg = F, pch = 19, cex = 0.2)#, main = "Annelida")
plotResiduals(simulationOutput_flat, quantreg = F, pch = 19, cex = 0.2)#, main = "Tricoptera")
plotResiduals(simulationOutput_caddis, quantreg = F, pch = 19, cex = 0.2)#, main = "Flatworms")
plotResiduals(simulationOutput_beet, quantreg = F, pch = 19, cex = 0.2)#, main = "Coleoptera")
plotResiduals(simulationOutput_dragon, quantreg = F, pch = 19, cex = 0.2)#, main = "Odonata")
plotResiduals(simulationOutput_may, quantreg = F, pch = 19, cex = 0.2)#, main = "Ephemeroptera")
```
Some deviations from statistical tests, but not enough to rule out the models. 



## Predictions 


As we have linear models with no interactions or polynomials we can just take arbitrary values for a baseline (e.g., mean values the variables in the "modern" dataset). We can calculate macroinvertebrate scores for this baseline. Then we change the baseline values for the explanatory variables by how much INCA predicts for each scenario. Then we calculate the macroinvertebrate scores again and measure the difference. We can plot the data as points (median INCA predictions) with error bars (INCA upper and lower percentiles). We are not making ***specific*** predictions about what the macroinvertebrate score will be at a specific site. Instead, we are just asking what the change will be due to changes in the fixed effects. So calculating the prediction intervals (using standard error of regression models) doesn't make much sense. If we wanted to predict a specific score, we would need to account for the standard error of the regression model coefficients and we would need to report the prediction interval. We would need to do this for the range of INCA results as well - so there would be considerable error propagation. Instead, we are just interested in change so we just need to carry through the INCA uncertainty to the forecast uncertainty. 




**Establish baseline**

```{r}
flow <- mean(glmm_data$flow) + 50 # preventing negative values
Temperature.of.Water_cel <- mean(glmm_data$Temperature.of.Water_cel) 
Oxygen..Dissolved.as.O2_mg.l <- mean(glmm_data$Oxygen..Dissolved.as.O2_mg.l) 
Orthophosphate..reactive.as.P_mg.l <- mean(glmm_data$Orthophosphate..reactive.as.P_mg.l)
SILT_CLAY <- mean(glmm_data$SILT_CLAY)
Ammonia.un.ionised.as.N_mg.l <- mean(glmm_data$Ammonia.un.ionised.as.N_mg.l) +5 # preventing negative values
Nitrate.as.N_mg.l <- mean(glmm_data$Nitrate.as.N_mg.l)

SITE_ID <- 33208 # random site ID

baseline <- data.frame(flow, Temperature.of.Water_cel,
                            Oxygen..Dissolved.as.O2_mg.l,
                            Orthophosphate..reactive.as.P_mg.l,
                            SILT_CLAY, Ammonia.un.ionised.as.N_mg.l,
                            Nitrate.as.N_mg.l, SITE_ID)

# Calculate macroinvertebrate scores for baselines 
baseline_aspt <- predict(aspt_model, baseline)
baseline_total <- predict(total_model, baseline)
baseline_rich <- predict(rich_model, baseline)
baseline_stone <- predict(plecoptera_model, baseline)
baseline_worms <- predict(annelid_model, baseline)
baseline_flat <- predict(flat_model, baseline)
baseline_beet <- predict(beet_model, baseline)
baseline_dragon <- predict(dragon_model, baseline)
baseline_may <- predict(mayfly_model, baseline)
baseline_trich <- predict(trichoptera_model, baseline)


```



**Load in scenarios** 

```{r}
scenarios <- read.csv("data/scenarios/scenarios.csv",
                      header = T)

# give a unique name to flow
scenarios$Variable <- gsub("flow", "Fl", scenarios$Variable)

```


**define a function for making the predictions**

```{r}

## Create a function that takes a climate, a land use and a site and returns
## .. the median, lower and upper change for all the macroinvertebrate responses 

macro_predictions <- function(location, climate, land) {
  
  data_test <- scenarios %>%
  filter(Site == location) %>%
  filter(RCP == climate) %>%
  filter(LandUse == land) 

  #### Median ####
  median <- data_test %>%
    filter(Estimate == "median") %>%
    select(-c(X, Site, RCP, LandUse, Estimate)) %>%
    pivot_wider(names_from = "Variable", values_from = "Differences") %>%
    # add in fixed baselines 
    mutate(SITE_ID = baseline$SITE_ID) %>%
    mutate(SILT_CLAY = baseline$SILT_CLAY) %>%
    # add in predicted changes due to INCA 
    mutate(flow = baseline$flow + Fl) %>%
    mutate(Temperature.of.Water_cel = baseline$Temperature.of.Water_cel + temp) %>%
    mutate(Oxygen..Dissolved.as.O2_mg.l = baseline$Oxygen..Dissolved.as.O2_mg.l + DO) %>%
    mutate(Orthophosphate..reactive.as.P_mg.l = baseline$Orthophosphate..reactive.as.P_mg.l + P) %>%
    mutate(Nitrate.as.N_mg.l = baseline$Nitrate.as.N_mg.l + Ni) %>%
    mutate(Ammonia.un.ionised.as.N_mg.l = baseline$Ammonia.un.ionised.as.N_mg.l + Am) 
  
  median_aspt <- predict(aspt_model, median) - baseline_aspt
  median_total <- predict(total_model, median) - baseline_total
  median_rich <- predict(rich_model, median) - baseline_rich
  median_stone <- predict(plecoptera_model, median) - baseline_stone
  median_worm <- predict(annelid_model, median) - baseline_worms
  median_flat <- predict(flat_model, median) - baseline_flat
  median_beet <- predict(beet_model, median) - baseline_beet
  median_dragon <- predict(dragon_model, median) - baseline_dragon
  median_may <- predict(mayfly_model, median) - baseline_may
  median_trich <- predict(trichoptera_model, median) - baseline_trich

  
  
  #### Lower ####
  lower <- data_test %>%
    filter(Estimate == "lower") %>%
    select(-c(X, Site, RCP, LandUse, Estimate)) %>%
    pivot_wider(names_from = "Variable", values_from = "Differences") %>%
    # add in fixed baselines 
    mutate(SITE_ID = baseline$SITE_ID) %>%
    mutate(SILT_CLAY = baseline$SILT_CLAY) %>%
    # add in predicted changes due to INCA 
    mutate(flow = baseline$flow + Fl) %>%
    mutate(Temperature.of.Water_cel = baseline$Temperature.of.Water_cel + temp) %>%
    mutate(Oxygen..Dissolved.as.O2_mg.l = baseline$Oxygen..Dissolved.as.O2_mg.l + DO) %>%
    mutate(Orthophosphate..reactive.as.P_mg.l = baseline$Orthophosphate..reactive.as.P_mg.l + P) %>%
    mutate(Nitrate.as.N_mg.l = baseline$Nitrate.as.N_mg.l + Ni) %>%
    mutate(Ammonia.un.ionised.as.N_mg.l = baseline$Ammonia.un.ionised.as.N_mg.l + Am) 
  
  lower_aspt <- predict(aspt_model, lower) - baseline_aspt
  lower_total <- predict(total_model, lower) - baseline_total
  lower_rich <- predict(rich_model, lower) - baseline_rich
  lower_stone <- predict(plecoptera_model, lower) - baseline_stone
  lower_worm <- predict(annelid_model, lower) - baseline_worms
  lower_flat <- predict(flat_model, lower) - baseline_flat
  lower_beet <- predict(beet_model, lower) - baseline_beet
  lower_dragon <- predict(dragon_model, lower) - baseline_dragon
  lower_may <- predict(mayfly_model, lower) - baseline_may
  lower_trich <- predict(trichoptera_model, lower) - baseline_trich
  
  #### Upper ####
  upper <- data_test %>%
    filter(Estimate == "upper") %>%
    select(-c(X, Site, RCP, LandUse, Estimate)) %>%
    pivot_wider(names_from = "Variable", values_from = "Differences") %>%
    # add in fixed baselines 
    mutate(SITE_ID = baseline$SITE_ID) %>%
    mutate(SILT_CLAY = baseline$SILT_CLAY) %>%
    # add in predicted changes due to INCA 
    mutate(flow = baseline$flow + Fl) %>%
    mutate(Temperature.of.Water_cel = baseline$Temperature.of.Water_cel + temp) %>%
    mutate(Oxygen..Dissolved.as.O2_mg.l = baseline$Oxygen..Dissolved.as.O2_mg.l + DO) %>%
    mutate(Orthophosphate..reactive.as.P_mg.l = baseline$Orthophosphate..reactive.as.P_mg.l + P) %>%
    mutate(Nitrate.as.N_mg.l = baseline$Nitrate.as.N_mg.l + Ni) %>%
    mutate(Ammonia.un.ionised.as.N_mg.l = baseline$Ammonia.un.ionised.as.N_mg.l + Am) 
  
  upper_aspt <- predict(aspt_model, upper) - baseline_aspt
  upper_total <- predict(total_model, upper) - baseline_total
  upper_rich <- predict(rich_model, upper) - baseline_rich
  upper_stone <- predict(plecoptera_model, upper) - baseline_stone
  upper_worm <- predict(annelid_model, upper) - baseline_worms
  upper_flat <- predict(flat_model, upper) - baseline_flat
  upper_beet <- predict(beet_model, upper) - baseline_beet
  upper_dragon <- predict(dragon_model, upper) - baseline_dragon
  upper_may <- predict(mayfly_model, upper) - baseline_may
  upper_trich <- predict(trichoptera_model, upper) - baseline_trich
  
  combined <- data.frame(median_aspt, median_total, median_rich,
                         lower_aspt, lower_total, lower_rich, 
                         upper_aspt, upper_total, upper_rich,
                         upper_stone, upper_worm, lower_stone, lower_worm, median_stone, median_worm,
                         median_flat, median_beet, median_dragon, median_may, median_trich,
                         upper_flat, upper_beet, upper_dragon, upper_may, upper_trich,
                         lower_flat, lower_beet, lower_dragon, lower_may, lower_trich)
  
  combined$scenario <- paste(location, climate, land,sep = ".")
  
  return(combined)
  
}


```


**Run scenarios** 

```{r}

C_26_no <- macro_predictions("Crick", "rcp26", "no_land")
O_26_no <- macro_predictions("Ox", "rcp26", "no_land")
L_26_no <- macro_predictions("Lon", "rcp26", "no_land")
C_45_no <- macro_predictions("Crick", "rcp45", "no_land")
O_45_no <- macro_predictions("Ox", "rcp45", "no_land")
L_45_no <- macro_predictions("Lon", "rcp45", "no_land")
E_26_no <- macro_predictions("Egham", "rcp26", "no_land")
E_45_no <- macro_predictions("Egham", "rcp45", "no_land")
Ca_26_no <- macro_predictions("Caver", "rcp26", "no_land")
Ca_45_no <- macro_predictions("Caver", "rcp45", "no_land")

C_26_ag <- macro_predictions("Crick", "rcp26", "agri")
O_26_ag <- macro_predictions("Ox", "rcp26", "agri")
L_26_ag <- macro_predictions("Lon", "rcp26", "agri")
C_45_ag <- macro_predictions("Crick", "rcp45", "agri")
O_45_ag <- macro_predictions("Ox", "rcp45", "agri")
L_45_ag <- macro_predictions("Lon", "rcp45", "agri")
E_26_ag <- macro_predictions("Egham", "rcp26", "agri")
E_45_ag <- macro_predictions("Egham", "rcp45", "agri")
Ca_26_ag <- macro_predictions("Caver", "rcp26", "agri")
Ca_45_ag <- macro_predictions("Caver", "rcp45", "agri")

C_26_agP <- macro_predictions("Crick", "rcp26", "agri_P")
O_26_agP <- macro_predictions("Ox", "rcp26", "agri_P")
L_26_agP <- macro_predictions("Lon", "rcp26", "agri_P")
C_45_agP <- macro_predictions("Crick", "rcp45", "agri_P")
O_45_agP <- macro_predictions("Ox", "rcp45", "agri_P")
L_45_agP <- macro_predictions("Lon", "rcp45", "agri_P")
E_26_agP <- macro_predictions("Egham", "rcp26", "agri_P")
E_45_agP <- macro_predictions("Egham", "rcp45", "agri_P")
Ca_26_agP <- macro_predictions("Caver", "rcp26", "agri_P")
Ca_45_agP <- macro_predictions("Caver", "rcp45", "agri_P")

macroinvertebrate_changes <- rbind(C_26_no, O_26_no, L_26_no, E_26_no, Ca_26_no,
                                   C_45_no, O_45_no, L_45_no, E_45_no, Ca_45_no,
                                   
                                   C_26_ag, O_26_ag, L_26_ag, E_26_ag, Ca_26_ag,
                                   C_45_ag, O_45_ag, L_45_ag, E_45_ag, Ca_45_ag,
                                   
                                   C_26_agP, O_26_agP, L_26_agP, E_26_agP, Ca_26_agP,
                                   C_45_agP, O_45_agP, L_45_agP, E_45_agP, Ca_45_agP)

rm(C_26_no, O_26_no, L_26_no, C_45_no, O_45_no, L_45_no, 
   C_26_ag, O_26_ag, L_26_ag, C_45_ag, O_45_ag, L_45_ag,
   C_26_agP, O_26_agP, L_26_agP, C_45_agP, O_45_agP, L_45_agP,
   E_26_no, Ca_26_no, E_45_no, Ca_45_no, E_26_ag, Ca_26_ag,
   E_45_ag, Ca_45_ag, E_26_agP, Ca_26_agP, E_45_agP, Ca_45_agP)

```

**Save results** 

```{r}
write.csv(macroinvertebrate_changes, "data/macro_results.csv")
```
























