---
title: "Plotting Results"
author: "James Orr"
output:
   html_notebook:
     code_folding: hide
     theme: flatly
     toc: true
     toc_depth: 4
     number_sections: no
---

This is a short notebook where we create the figures for the paper. 

**Set up Environment**

```{r, echo=TRUE, results='hide'}
#### Clear  environment 
rm(list = ls())         

#### Required packages
library(tidyverse)        # for general use
library(sf)               # classes and functions for vector data
library(raster)           # classes and functions for raster data
library(nngeo)            # for the sf_remove_holes function
library(ggspatial)        # for scale bar and north arrow 
library(readxl)           # reading excel files
```


## Map of catchment 

**Load and organise data**

```{r}
######### Thames Catchment Boundary ############
catchment_shape <- st_read("data/geo/catch/WFD_River_Basin_Districts_Cycle_2.shp") %>%
  # subset the Thames catchment 
  filter(rbd_name == "Thames")


### Sub-catchments ###
sub_catchments_shape <- st_read("data/geo/catch/WFD_Surface_Water_Management_Catchments_Cycle_3.shp") %>%
  filter(!MNCAT_NAME %in% c("Medway", "Thames TraC", "Kent North", "London",
                            "Darent and Cray", "Lee Upper", "Roding Beam and Ingrebourne",
                            "Essex South"))
  
catchment_main <- st_remove_holes(st_union(sub_catchments_shape)) 

######### River shapefile ############
rivers_shape <- st_read("data/geo/rivers/data/WatercourseLink.shp") %>%
  # reduce the "resolution" of the shapefile 
  #st_simplify(dTolerance = 100) %>%
  filter(st_within(x = ., y = catchment_main, sparse = FALSE))


#### Thames main river #######
thames <- st_read("data/geo/RBD_Thames_shapefile/WFD_River_Water_Bodies_Cycle_3.shp") %>%
  filter(WB_NAME %in% c("Thames (Churn to Coln)",
                        "Thames (Cookham to Egham)",
                        "Thames (Egham to Teddington)",
                        "Thames (Evenlode to Thame)",
                        "Thames (Kemble to Waterhay Bridge)",
                        "Thames (Leach to Evenlode)",
                        "Thames (Reading to Cookham)",
                        "Thames (Waterhaybridge to Cricklade) and Chelworth Brook",
                        "Thames Wallingford to Caversham"))


######### UK Boundaries ############ (http://www.diva-gis.org/gdata)
uk_admin <- st_read("data/geo/GBR_adm/GBR_adm0.shp") %>%
  st_simplify(dTolerance = 100)


######## Elevation data for Thames catchment #########
elevation <- raster("data/geo/GBR_alt/GBR_alt.grd") 

new_crs <- new_crs <- CRS("+init=epsg:27700")       # WKT formate
elevation <- projectRaster(elevation, crs=new_crs)

elevation <- mask(elevation, sub_catchments_shape)

elevation_df <- as.data.frame(rasterToPoints(elevation))


#### manually locating three sites #### 
easting = c(410306, 452409, 515411, 471394, 503607)
northing = c(193909, 202855, 168473, 174426, 170277)
place = c("Cricklade", "Oxford", "London", "Caversham", "Egham")
river_sites <- data.frame(easting,  northing, place)

river_sites <- st_as_sf(river_sites, coords = c("easting", "northing")) %>%
  st_set_crs(., 27700)


### London ###
london <- st_read("data/geo/london/London_Borough_Excluding_MHW.shp") %>%
  st_set_crs(., 27700)
london<-st_union(london)

```



```{r}
### Invert data ### 
macro_data <- list.files(path = "data/combined/",
                  pattern="*.csv",
                  full.names = T) %>%
  map_df(~read_csv(., col_types = cols(.default = "c")))

### Need to filter in the same way I do for my regressions 
macro_data <- macro_data %>%
  
  dplyr::select(c(
    
          # general descriptors of each sample 
          SITE_ID, SAMPLE_DATE_YEAR, SAMPLE_DATE_MONTH,
          
          # habitat quality variables
          DISCHARGE, WIDTH, SLOPE, 
          SILT_CLAY, BOULDERS_COBBLES, PEBBLES_GRAVEL, SAND,
          
          # water quality variables (that can be predicted by INCA)
          `Temperature of Water_cel`,
          `Oxygen, Dissolved as O2_mg/l`,
          `Ammonia un-ionised as N_mg/l`, 
          `Nitrate as N_mg/l`,
          `Orthophosphate, reactive as P_mg/l`, 
           
          # response variables 
          WHPT_ASPT, WHPT_TOTAL, 
          macro_richness, 
          mollusc, 
          total_abundance,
          ept_richness,
          ept_abundance,
          `insect - true fly (Diptera)`,
          `insect - stonefly (Plecoptera)`,
          `insect - mayfly (Ephemeroptera)`,
          `insect - dragonfly (Odonata)`,
          `insect - caddis fly (Trichoptera)`,
          `insect - beetle (Coleoptera)`,
          `flatworm (Turbellaria)`,
          crustacean,
          annelid)) %>%
  
  drop_na()  # llmer requires full cases 

# Need to go back and grab the full eastings and full northings 
data_site <- read.csv("data/eco/inverts/INV_OPEN_DATA_SITE_2022-02-01.csv", 
                     header = T)

data_site <- data_site %>%
  filter(SITE_ID %in% macro_data$SITE_ID) %>%
  dplyr::select(c(SITE_ID, FULL_EASTING, FULL_NORTHING)) 

data_site$SITE_ID <- as.character(data_site$SITE_ID)

macro_data <- left_join(macro_data, data_site, by = "SITE_ID")

macro_data <- st_as_sf(macro_data, coords = c("FULL_EASTING", 
                                              "FULL_NORTHING")) %>%
  st_set_crs(., 27700) %>%
  filter(st_within(x = ., y = catchment_main, sparse = FALSE))
```

```{r}
ggplot(data = uk_admin) +
  
  geom_sf(fill = rgb(0.8, 0.8, 0.8), color = NA) +
  
  geom_sf(data = catchment_main, fill = "#a18660", color = NA) +
  
  theme_void()
  
```


```{r, fig.width=8, fig.height=4.8}
# colour ramp
elevation_colours <- colorRampPalette(c("#99d38b",
                                        "#fffabe",
                                        "#ffdc8f",
                                        "#a18660"))
theme_set(theme_void())

ggplot() +
  
  # elevation
  geom_raster(data = elevation_df, aes(x = x, y = y, fill = GBR_alt)) +
  scale_fill_gradientn(colors = elevation_colours(20), name = "Elevation (m)") +

  # rivers
  geom_sf(data = rivers_shape, color = rgb(0.1, 0.1, 1, 0.2)) +
  
  # Relevant-sub catchment 
  geom_sf(data = catchment_main, color = rgb(0.5, 0.5, 0.5), linewidth = 0.5, fill =NA) +
  
  # Scale bar and north arrow 
  annotation_scale(location = "bl", line_width = 2, text_cex = 1.2,
                   pad_x = unit(1.6, "in"),
                   pad_y = unit(1, "in")) +
  annotation_north_arrow(location = "bl",
                         pad_x = unit(1.6-0.85, "in"),
                         pad_y = unit(1-0.2, "in")) +
  
  # Main Thames 
  geom_sf(data = thames, color = rgb(0.2, 0.4, 1, 0.3), linewidth = 2.5) +
  
  # invert sites
  geom_sf(data = macro_data, color = rgb(0.3, 0.3, 0.3), size = 1.5) +
  
  # river sites 
  geom_sf(data = river_sites, color = "red") +

  # Theme adjustments
  theme(legend.direction="horizontal", 
        legend.position = "top", 
        legend.justification = "right",
        legend.text=element_text(size = 12),
        legend.title=element_text(size =  15))


```


## Water quality forecasts 

```{r}
#### No land-use change ####
no_land_median <- read_excel("data/inca/summary_2080s.xlsx",
                             sheet = "median_month")

no_land_lower <- read_excel("data/inca/summary_2080s.xlsx",
                             sheet = "lower_month")

no_land_upper <- read_excel("data/inca/summary_2080s.xlsx",
                             sheet = "upper_month")

#### Agricultural intensification ####

agri_median <- read_excel("data/inca/summary_2080s_agri.xlsx",
                             sheet = "median_month")

agri_lower <- read_excel("data/inca/summary_2080s_agri.xlsx",
                             sheet = "lower_month")

agri_upper <- read_excel("data/inca/summary_2080s_agri.xlsx",
                             sheet = "upper_month")

#### Agricultural intensification + P removal ####

agri_P_median <- read_excel("data/inca/summary_2080s_agri_P.xlsx",
                             sheet = "median_month")

agri_P_lower <- read_excel("data/inca/summary_2080s_agri_P.xlsx",
                             sheet = "lower_month")

agri_P_upper <- read_excel("data/inca/summary_2080s_agri_P.xlsx",
                             sheet = "upper_month")

```

```{r}
data_cleaning <- function(dataset, name) {

  dataset <- dataset %>%
    dplyr::select(c(...1,
                    baseline_TR01_flow,
                    baseline_TR01_watertemperature,
                    baseline_TR01_watercolSRP,
                    baseline_TR01_dissolvedoxygen,
                    baseline_TR01_ammonium,
                    baseline_TR01_nitrate,
                    rcp45_TR01_flow,
                    rcp45_TR01_watertemperature,
                    rcp45_TR01_watercolSRP,
                    rcp45_TR01_dissolvedoxygen,
                    rcp45_TR01_ammonium,
                    rcp45_TR01_nitrate,
                    
                    baseline_TR22_flow,
                    baseline_TR22_watertemperature,
                    baseline_TR22_watercolSRP,
                    baseline_TR22_dissolvedoxygen,
                    baseline_TR22_ammonium,
                    baseline_TR22_nitrate,
                    rcp45_TR22_flow,
                    rcp45_TR22_watertemperature,
                    rcp45_TR22_watercolSRP,
                    rcp45_TR22_dissolvedoxygen,
                    rcp45_TR22_ammonium,
                    rcp45_TR22_nitrate)) %>%
    mutate(scenario = name)
  
  return(dataset)
  
}

```

```{r}
no_land_median_short <- data_cleaning(no_land_median, "no_land_median")
no_land_lower_short <- data_cleaning(no_land_lower, "no_land_lower")
no_land_upper_short <- data_cleaning(no_land_upper, "no_land_upper")
agri_P_median_short <- data_cleaning(agri_P_median, "agri_P_median")
agri_P_lower_short <- data_cleaning(agri_P_lower, "agri_P_lower")
agri_P_upper_short <- data_cleaning(agri_P_upper, "agri_P_upper")
agri_median_short <- data_cleaning(agri_median, "agri_median")
agri_lower_short <- data_cleaning(agri_lower, "agri_lower")
agri_upper_short <- data_cleaning(agri_upper, "agri_upper")

no_land <- rbind(no_land_median_short, no_land_lower_short, no_land_upper_short)
agri <- rbind(agri_median_short, agri_lower_short, agri_upper_short)
agri_P <- rbind(agri_P_median_short, agri_P_lower_short, agri_P_upper_short)


no_land_wider <- pivot_wider(no_land, 
                          names_from = scenario,
                          values_from = c(baseline_TR01_flow,
                    baseline_TR01_watertemperature,
                    baseline_TR01_watercolSRP,
                    baseline_TR01_dissolvedoxygen,
                    baseline_TR01_ammonium,
                    baseline_TR01_nitrate,
                    rcp45_TR01_flow,
                    rcp45_TR01_watertemperature,
                    rcp45_TR01_watercolSRP,
                    rcp45_TR01_dissolvedoxygen,
                    rcp45_TR01_ammonium,
                    rcp45_TR01_nitrate,
                    
                    baseline_TR22_flow,
                    baseline_TR22_watertemperature,
                    baseline_TR22_watercolSRP,
                    baseline_TR22_dissolvedoxygen,
                    baseline_TR22_ammonium,
                    baseline_TR22_nitrate,
                    rcp45_TR22_flow,
                    rcp45_TR22_watertemperature,
                    rcp45_TR22_watercolSRP,
                    rcp45_TR22_dissolvedoxygen,
                    rcp45_TR22_ammonium,
                    rcp45_TR22_nitrate))

agri_P_wider <- pivot_wider(agri_P, 
                          names_from = scenario,
                          values_from = c(baseline_TR01_flow,
                    baseline_TR01_watertemperature,
                    baseline_TR01_watercolSRP,
                    baseline_TR01_dissolvedoxygen,
                    baseline_TR01_ammonium,
                    baseline_TR01_nitrate,
                    rcp45_TR01_flow,
                    rcp45_TR01_watertemperature,
                    rcp45_TR01_watercolSRP,
                    rcp45_TR01_dissolvedoxygen,
                    rcp45_TR01_ammonium,
                    rcp45_TR01_nitrate,
                    
                    baseline_TR22_flow,
                    baseline_TR22_watertemperature,
                    baseline_TR22_watercolSRP,
                    baseline_TR22_dissolvedoxygen,
                    baseline_TR22_ammonium,
                    baseline_TR22_nitrate,
                    rcp45_TR22_flow,
                    rcp45_TR22_watertemperature,
                    rcp45_TR22_watercolSRP,
                    rcp45_TR22_dissolvedoxygen,
                    rcp45_TR22_ammonium,
                    rcp45_TR22_nitrate))

agri_wider <- pivot_wider(agri, 
                          names_from = scenario,
                          values_from = c(baseline_TR01_flow,
                    baseline_TR01_watertemperature,
                    baseline_TR01_watercolSRP,
                    baseline_TR01_dissolvedoxygen,
                    baseline_TR01_ammonium,
                    baseline_TR01_nitrate,
                    rcp45_TR01_flow,
                    rcp45_TR01_watertemperature,
                    rcp45_TR01_watercolSRP,
                    rcp45_TR01_dissolvedoxygen,
                    rcp45_TR01_ammonium,
                    rcp45_TR01_nitrate,
                    
                    baseline_TR22_flow,
                    baseline_TR22_watertemperature,
                    baseline_TR22_watercolSRP,
                    baseline_TR22_dissolvedoxygen,
                    baseline_TR22_ammonium,
                    baseline_TR22_nitrate,
                    rcp45_TR22_flow,
                    rcp45_TR22_watertemperature,
                    rcp45_TR22_watercolSRP,
                    rcp45_TR22_dissolvedoxygen,
                    rcp45_TR22_ammonium,
                    rcp45_TR22_nitrate))

```

Crickdale 

```{r, fig.height=2, fig.width=4.25, dpi = 300}

par(mfrow = c(2, 3),
    mar=c(3, 3, 3, 3)) 

###### Flow, baseline, rcp45 all at cricklade 
plot(no_land_wider$...1, no_land_wider$baseline_TR01_flow_no_land_median,
     type = "l", xlab = "", ylab = "Flow",
     ylim = c(1, 12.5), las = 1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR01_flow_no_land_lower,
          rev(no_land_wider$rcp45_TR01_flow_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR01_flow_no_land_median, col = rgb(255/255, 177/255, 177/255), lwd = 2)



###### Temperature, baseline, rcp45 all at cricklade 
plot(no_land_wider$...1, no_land_wider$baseline_TR01_watertemperature_no_land_median,
     type = "l", xlab = "", ylab = "Water Temperature",
     ylim = c(3, 22), las = 1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR01_watertemperature_no_land_lower,
          rev(no_land_wider$rcp45_TR01_watertemperature_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR01_watertemperature_no_land_median, col = rgb(255/255, 177/255, 177/255), lwd = 2)



###### Oxygen, baseline, rcp45 all at cricklade 
plot(no_land_wider$...1, no_land_wider$baseline_TR01_dissolvedoxygen_no_land_median,
     type = "l", xlab = "", ylab = "Dissolved Oxygen",
     ylim = c(8, 13), las =1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR01_dissolvedoxygen_no_land_lower,
          rev(no_land_wider$rcp45_TR01_dissolvedoxygen_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR01_dissolvedoxygen_no_land_median, col = rgb(255/255, 177/255, 177/255), lwd = 2)


###### P, base v agri_P, rcp45, cricklade
plot(no_land_wider$...1, no_land_wider$baseline_TR01_watercolSRP_no_land_median,
     type = "l", xlab = "", ylab = "Phosphorus",
     ylim = c(0, 0.7), las = 1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR01_watercolSRP_no_land_lower,
          rev(no_land_wider$rcp45_TR01_watercolSRP_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR01_watercolSRP_no_land_median, col = rgb(1, 0, 0, 0.3), lwd = 2)

polygon(c(agri_P_wider$...1,
          rev(agri_P_wider$...1)),
        c(agri_P_wider$rcp45_TR01_watercolSRP_agri_P_lower,
          rev(agri_P_wider$rcp45_TR01_watercolSRP_agri_P_upper)),col = rgb(0, 0, 1, 0.1), border = FALSE)
lines(agri_P_wider$...1, agri_P_wider$rcp45_TR01_watercolSRP_agri_P_median, col = rgb(0, 0, 1, 0.3), lwd = 2, lty = 3)


###### Nitrate, base v agri, rcp45, cricklade
plot(no_land_wider$...1, no_land_wider$baseline_TR01_nitrate_no_land_median,
     type = "l", xlab = "", ylab = "Nitrate",
     ylim = c(6, 7.7), las = 1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR01_nitrate_no_land_lower,
          rev(no_land_wider$rcp45_TR01_nitrate_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR01_nitrate_no_land_median, col = rgb(1, 0, 0, 0.3), lwd = 2)

polygon(c(agri_wider$...1,
          rev(agri_wider$...1)),
        c(agri_wider$rcp45_TR01_nitrate_agri_lower,
          rev(agri_wider$rcp45_TR01_nitrate_agri_upper)),col = rgb(1, 0, 0, 0.3), border = FALSE)
lines(agri_wider$...1, agri_wider$rcp45_TR01_nitrate_agri_median, col = rgb(1, 0, 0, 0.5), lwd = 2, lty = 2)


###### Ammonium, base v agri, rcp45, cricklade
plot(no_land_wider$...1, no_land_wider$baseline_TR01_ammonium_no_land_median,
     type = "l", xlab = "", ylab = "Ammonium",
     ylim = c(0.035, 0.135), las = 1, lty = 2)

polygon(c(agri_wider$...1,
          rev(agri_wider$...1)),
        c(agri_wider$rcp45_TR01_ammonium_agri_lower,
          rev(agri_wider$rcp45_TR01_ammonium_agri_upper)),col = rgb(1, 0, 0, 0.3), border = FALSE)
lines(agri_wider$...1, agri_wider$rcp45_TR01_ammonium_agri_median, col = rgb(1, 0, 0, 0.5), lwd = 2, lty = 2)

```

London 

```{r, fig.height=2, fig.width=4.25, dpi = 300}

par(mfrow = c(2, 3),
    mar=c(3, 3, 3, 3)) 

###### Flow, baseline, rcp45 all at London 
plot(no_land_wider$...1, no_land_wider$baseline_TR22_flow_no_land_median,
     type = "l", xlab = "", ylab = "Flow",
     ylim = c(0, 230), 
     las = 1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR22_flow_no_land_lower,
          rev(no_land_wider$rcp45_TR22_flow_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR22_flow_no_land_median, col = rgb(255/255, 177/255, 177/255), lwd = 2)



###### Temperature, baseline, rcp45 all at London 
plot(no_land_wider$...1, no_land_wider$baseline_TR22_watertemperature_no_land_median,
     type = "l", xlab = "", ylab = "Water Temperature",
     ylim = c(3, 22), las = 1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR22_watertemperature_no_land_lower,
          rev(no_land_wider$rcp45_TR22_watertemperature_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR22_watertemperature_no_land_median, col = rgb(255/255, 177/255, 177/255), lwd = 2)



###### Oxygen, baseline, rcp45 all at London 
plot(no_land_wider$...1, no_land_wider$baseline_TR22_dissolvedoxygen_no_land_median,
     type = "l", xlab = "", ylab = "Dissolved Oxygen",
     ylim = c(7, 12.5), las =1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR22_dissolvedoxygen_no_land_lower,
          rev(no_land_wider$rcp45_TR22_dissolvedoxygen_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR22_dissolvedoxygen_no_land_median, col = rgb(255/255, 177/255, 177/255), lwd = 2)


###### P, base v agri_P, rcp45, London
plot(no_land_wider$...1, no_land_wider$baseline_TR22_watercolSRP_no_land_median,
     type = "l", xlab = "", ylab = "Phosphorus",
     ylim = c(0, 0.75), las = 1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR22_watercolSRP_no_land_lower,
          rev(no_land_wider$rcp45_TR22_watercolSRP_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR22_watercolSRP_no_land_median, col = rgb(1, 0, 0, 0.3), lwd = 2)

polygon(c(agri_P_wider$...1,
          rev(agri_P_wider$...1)),
        c(agri_P_wider$rcp45_TR22_watercolSRP_agri_P_lower,
          rev(agri_P_wider$rcp45_TR22_watercolSRP_agri_P_upper)),col = rgb(0, 0, 1, 0.1), border = FALSE)
lines(agri_P_wider$...1, agri_P_wider$rcp45_TR22_watercolSRP_agri_P_median, col = rgb(0, 0, 1, 0.3), lwd = 2, lty = 3)


###### Nitrate, base v agri, rcp45, London
plot(no_land_wider$...1, no_land_wider$baseline_TR22_nitrate_no_land_median,
     type = "l", xlab = "", ylab = "Nitrate",
     ylim = c(5.8, 7.5), las = 1, lty = 2)
polygon(c(no_land_wider$...1,
          rev(no_land_wider$...1)),
        c(no_land_wider$rcp45_TR22_nitrate_no_land_lower,
          rev(no_land_wider$rcp45_TR22_nitrate_no_land_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
lines(no_land_wider$...1, no_land_wider$rcp45_TR22_nitrate_no_land_median, col = rgb(1, 0, 0, 0.3), lwd = 2)

polygon(c(agri_wider$...1,
          rev(agri_wider$...1)),
        c(agri_wider$rcp45_TR22_nitrate_agri_lower,
          rev(agri_wider$rcp45_TR22_nitrate_agri_upper)),col = rgb(1, 0, 0, 0.3), border = FALSE)
lines(agri_wider$...1, agri_wider$rcp45_TR22_nitrate_agri_median, col = rgb(1, 0, 0, 0.5), lwd = 2, lty = 2)


###### Ammonium, base v agri, rcp45, London
plot(no_land_wider$...1, no_land_wider$baseline_TR22_ammonium_no_land_median,
     type = "l", xlab = "", ylab = "Ammonium",
     ylim = c(0.07, 0.15), las = 1, lty = 2)

polygon(c(agri_wider$...1,
          rev(agri_wider$...1)),
        c(agri_wider$rcp45_TR22_ammonium_agri_lower,
          rev(agri_wider$rcp45_TR22_ammonium_agri_upper)),col = rgb(1, 0, 0, 0.3), border = FALSE)
lines(agri_wider$...1, agri_wider$rcp45_TR22_ammonium_agri_median, col = rgb(1, 0, 0, 0.5), lwd = 2, lty = 2)

```



```{r}
climate <- read.csv("data/inca/climate_change.csv", header = T)
```

```{r, fig.height=3, fig.width=4, dpi = 300}

par(mfrow = c(2, 2),
    mar=c(2, 1.5, 2, 1.5),   # plot margins
    oma=c(3, 3, 3, 3))   # outer margins 
  
plot(climate$Month, climate$rcp26_percipitation_median,
     type = "l", xlab = "", ylab = "Percipitation (% change)",
     ylim = c(-50, 50), las = 1, lty = 1,
     col = rgb(1, 0, 0, 0.3),
     lwd = 3)
polygon(c(climate$Month,
          rev(climate$Month)),
        c(climate$rcp26_percipitation_lower,
          rev(climate$rcp26_percipitation_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
abline(h = 0, lwd = 2, lty = 2, col = "grey60")

plot(climate$Month, climate$rcp45_percipitation_median,
     type = "l", xlab = "", ylab = "",
     ylim = c(-50, 50), las = 1, lty = 1,
     col = rgb(1, 0, 0, 0.5),
     lwd = 3)
polygon(c(climate$Month,
          rev(climate$Month)),
        c(climate$rcp45_percipitation_lower,
          rev(climate$rcp45_percipitation_upper)),col = rgb(1, 0, 0, 0.2), border = FALSE)
abline(h = 0, lwd = 2, lty = 2, col = "grey60")





plot(climate$Month, climate$rcp26_temperature_median,
     type = "l", xlab = "", ylab = "Temperature (% change)",
     ylim = c(-6.5, 6.5), las = 1, lty = 1,
     col = rgb(1, 0, 0, 0.3),
     lwd = 3)
polygon(c(climate$Month,
          rev(climate$Month)),
        c(climate$rcp26_temperature_lower,
          rev(climate$rcp26_temperature_upper)),col = rgb(1, 0, 0, 0.1), border = FALSE)
abline(h = 0, lwd = 2, lty = 2, col = "grey60")

plot(climate$Month, climate$rcp45_temperature_median,
     type = "l", xlab = "", ylab = "",
     ylim = c(-6.5, 6.5), las = 1, lty = 1,
     col = rgb(1, 0, 0, 0.5),
     lwd = 3)
polygon(c(climate$Month,
          rev(climate$Month)),
        c(climate$rcp45_temperature_lower,
          rev(climate$rcp45_temperature_upper)),col = rgb(1, 0, 0, 0.2), border = FALSE)
abline(h = 0, lwd = 2, lty = 2, col = "grey60")


```



## Macrointertebrate forecasts

**Load and organise data**

```{r}
macro_results <- read.csv("data/macro_results.csv", header = T)

macro_results <- macro_results %>%
  separate(scenario, into = c("Site", "Climate", "Land"), "\\.") %>%
  mutate(site_rank = ifelse(Site == "Crick", 1, 
                            ifelse(Site == "Ox", 2, 
                                   ifelse(Site == "Caver", 3, 
                                          ifelse(Site == "Egham", 4,5)))))

macro_results$site_rank <- as.numeric(macro_results$site_rank)
```



**Functions for figures** 

```{r}

individual_plot <- function(ymax, site_rank, median, upper, lower, yax, col) {
  
  plot(site_rank, median,
     xlim = c(0.25, 5.75), 
     ylim = c(-ymax, ymax),
     las = 1, xaxt='n', yaxt='n', xlab = "", ylab = "",
     cex = 0,
     frame.plot = TRUE)
  arrows(site_rank, upper,
         site_rank, lower,
         code = 0, lwd = 2, col = rgb(0.5, 0.5, 0.5))
  abline(h = 0, lty = 2)
  points(site_rank, median,
         cex = 2.5, pch = 21, col = rgb(0.15, 0.15, 0.15), bg = col)
  text(site_rank, median, labels = site_rank)
  
  if (yax == "yes") {
    axis(side = 2, las = 1)
  }
}


plot_score <- function(site_rank, median, upper, lower){
  
  # set y axis limits 
  ymax <- max(abs(upper), abs(lower))
  
  # set a color ramp 
  rbPal <- colorRampPalette(c(rgb(1, 0.4, 0.4), 'white', rgb(0.4, 0.4, 1)))
  breakpoints <- c(seq(from = -ymax, to = ymax, length.out = 100))
  macro_results$Col <- rbPal(length(breakpoints))[as.numeric(cut(median,breaks = breakpoints))]

  
  par(mfrow = c(2, 3),
      mar=c(0.2, 0.2, 0.2, 0.2),   # plot margins
      oma=c(3, 3, 3, 3))   # outer margins 
  
  individual_plot(ymax, 
                  site_rank[macro_results$Land == "no_land" & macro_results$Climate == "rcp26"],  
                  median[macro_results$Land == "no_land" & macro_results$Climate == "rcp26"], 
                  upper[macro_results$Land == "no_land" & macro_results$Climate == "rcp26"], 
                  lower[macro_results$Land == "no_land" & macro_results$Climate == "rcp26"], 
                  "yes", macro_results$Col[macro_results$Land == "no_land" & macro_results$Climate == "rcp26"])
  
    individual_plot(ymax, 
                  site_rank[macro_results$Land == "agri" & macro_results$Climate == "rcp26"],  
                  median[macro_results$Land == "agri" & macro_results$Climate == "rcp26"], 
                  upper[macro_results$Land == "agri" & macro_results$Climate == "rcp26"], 
                  lower[macro_results$Land == "agri" & macro_results$Climate == "rcp26"],
                  "no", macro_results$Col[macro_results$Land == "agri" & macro_results$Climate == "rcp26"])
    
    individual_plot(ymax, 
                  site_rank[macro_results$Land == "agri_P" & macro_results$Climate == "rcp26"],  
                  median[macro_results$Land == "agri_P" & macro_results$Climate == "rcp26"], 
                  upper[macro_results$Land == "agri_P" & macro_results$Climate == "rcp26"], 
                  lower[macro_results$Land == "agri_P" & macro_results$Climate == "rcp26"],
                  "no", macro_results$Col[macro_results$Land == "agri_P" & macro_results$Climate == "rcp26"])
      
    individual_plot(ymax, 
                  site_rank[macro_results$Land == "no_land" & macro_results$Climate == "rcp45"],  
                  median[macro_results$Land == "no_land" & macro_results$Climate == "rcp45"], 
                  upper[macro_results$Land == "no_land" & macro_results$Climate == "rcp45"], 
                  lower[macro_results$Land == "no_land" & macro_results$Climate == "rcp45"],
                  "yes", macro_results$Col[macro_results$Land == "no_land" & macro_results$Climate == "rcp45"])
  
    individual_plot(ymax, 
                  site_rank[macro_results$Land == "agri" & macro_results$Climate == "rcp45"],  
                  median[macro_results$Land == "agri" & macro_results$Climate == "rcp45"], 
                  upper[macro_results$Land == "agri" & macro_results$Climate == "rcp45"], 
                  lower[macro_results$Land == "agri" & macro_results$Climate == "rcp45"],
                  "no", macro_results$Col[macro_results$Land == "agri" & macro_results$Climate == "rcp45"])
    
    individual_plot(ymax, 
                  site_rank[macro_results$Land == "agri_P" & macro_results$Climate == "rcp45"],  
                  median[macro_results$Land == "agri_P" & macro_results$Climate == "rcp45"], 
                  upper[macro_results$Land == "agri_P" & macro_results$Climate == "rcp45"], 
                  lower[macro_results$Land == "agri_P" & macro_results$Climate == "rcp45"],
                  "no", macro_results$Col[macro_results$Land == "agri_P" & macro_results$Climate == "rcp45"])
      
      
}




```


WHPT_ASPT

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_aspt,
           macro_results$upper_aspt, 
           macro_results$lower_aspt)

```
WHPT_Total

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_total,
           macro_results$upper_total, 
           macro_results$lower_total)
```


Macro Richness

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_rich,
           macro_results$upper_rich, 
           macro_results$lower_rich)
```


Plecoptera 

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_stone,
           macro_results$upper_stone, 
           macro_results$lower_stone)
```


Annelida 

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_worm,
           macro_results$upper_worm, 
           macro_results$lower_worm)
```



Coleoptera 

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_beet,
           macro_results$upper_beet, 
           macro_results$lower_beet)
```



Flatworm 

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_flat,
           macro_results$upper_flat, 
           macro_results$lower_flat)
```



Ephemeroptera 

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_may,
           macro_results$upper_may, 
           macro_results$lower_may)
```


Trichoptera 

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_trich,
           macro_results$upper_trich, 
           macro_results$lower_trich)
```
Odonata 

```{r, fig.height=2.5, fig.width=2, dpi = 300}

plot_score(macro_results$site_rank,
           macro_results$median_dragon,
           macro_results$upper_dragon, 
           macro_results$lower_dragon)
```
